---
title: "Programming with Data - Defining Functions"
semester: Fall 2023
format:
  pdf:
    template: ../assignment-template.tex
editor: source
engine: knitr
---

When creating functions, it's often useful to keep the function definitions in a separate R script file (with file extension *.R) to avoid cluttering your Quarto document. For this assignment, create a separate R script file named `my_functions.R` and store all your function definitions in that file. Near the beginning of your Quarto document, include a code chunk like the following:

````{verbatim}
```{r}
#| include: false
source("my_functions.R")
```
````

This code chunk will load all of the functions defined in the file `my_functions.R` and you will be able to use these functions in any code chunks that come later in the document.

To demonstrate that your functions work, insert code chunks within your Quarto document that call your functions with the example values provided in the assignment below.

## Bushels per acre to kilograms per hectare


```{r}
#| include: false
source("my_functions.R")
```

In this assignment, we will create a set of functions for converting wheat yield from units in bushels per acre to kilograms per hectare.

1. Create a function named `bu_to_lbs` that has one argument for a value in units of bushels (bu) and converts it to units of pounds (lbs) using the following conversion factor: 1 bu wheat = 60 lbs wheat

```{r}
bu_to_lbs(1)
bu_to_lbs(30)
```

2. Create a function called `lbs_to_kg` that has one argument in pounds (lbs) and converts lbs to kilograms (kg) using a conversion factor of: 1 lbs = 0.453592 kg

```{r}
lbs_to_kg(2.2)
lbs_to_kg(60)
```

3. Create a function named `per_ac_to_per_ha` that has one argument and converts units on a per acre (ac) basis to units on a per hectare (ha) basis using the following conversion factor: 1 ac = 0.404686 ha *Note: Remember to divide by the conversion factor in this case because both are expressed on a per ac or ha basis i.e.* $\dfrac{1}{1 ac} = \dfrac{1}{0.404686 ha}$

```{r}
per_ac_to_per_ha(1)
per_ac_to_per_ha(1800)
```

4. Create a function named `bu_ac_to_kg_ha` that has one argument in bushels per acre and uses the previous three functions within a pipeline to perform the units conversion of wheat yield from bushels per acre to kilograms per hectare.

```{r}
bu_ac_to_kg_ha(30)
bu_ac_to_kg_ha(90)
```

