---
title: "Wrangling Data - Metadata and Data Import"
author: |
  | Data Science for Agriculture and Natural Resources
  | PLNT 5413 - Fall `r format(Sys.Date(),'%Y')`
output: pdf_document
link-citations: true
urlcolor: blue
margin: 1in
fontsize: 12pt
geometry: "left=1in,right=1in,top=1in,bottom=1.25in"
header-includes:
- \usepackage{titling}
- \usepackage{lastpage}
- \usepackage{fancyhdr}
- \newcommand{\aspace}[1]{\vspace{#1\baselineskip}}
- \renewcommand{\headrulewidth}{0pt}
- \fancypagestyle{plain}{ \fancyhf{} \fancyfoot[C]{\aspace{2}Page \thepage~of~\pageref{LastPage}} }
- \pagestyle{fancy}
- \fancyhf{}
- \fancyfoot[C]{\small \thetitle \\ Data Science for Agriculture and Natural Resources\\PLNT 5413 - Fall `r format(Sys.Date(),'%Y')`\\ Page \thepage~of~\pageref{LastPage}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

Estimating the total amount of biomass (the total mass of all individuals) in forests is important for understanding the global carbon budget and how the earth will respond to increases in carbon dioxide emissions. Measuring the mass of whole trees is a major effort and requires destructive harvest of the tree. Fortunately, we can estimate the mass of a tree based on its diameter.

We’re going to estimate the total tree biomass for trees in a 96 hectare area of the Western Ghats in India. The [raw data](https://canvas.okstate.edu/courses/109158/files/12019188) are available on [Ecological Archives](https://esapubs.org/Archive/). Unfortunately, the data are stored in a poor database structure and using all of the tree stems would be difficult without first tidying up the data. You will need to read the [metadata](https://esapubs.org/archive/ecol/E091/216/metadata.htm) to get familiar with the structure of the dataset.

1. After reading through the metadata, comment (2-3 sentences) on the completeness of the documentation. Compare this to your own documentation of metadata in your `README.txt` file from the `Organizing Data in Spreadsheets` assignment.

2. Describe at least 2 ways this dataset violates some of the best practices (e.g. tidy data, [Format Data](https://datacarpentry.org/spreadsheet-ecology-lesson/01-format-data/index.html), [Common Mistakes](https://datacarpentry.org/spreadsheet-ecology-lesson/02-common-mistakes/index.html)) for organizing data in spreadsheets.

3. Import the raw data into R.

```{r readdata, message = FALSE}

# Problem 3
tree_raw <- read_tsv('data/Macroplot_data_Rev.txt')

```

```{r answer}
tree_raw

```