---
title: 'Exploring Data - Visualization'
author: |
  | Data Science for Agriculture and Natural Resources
  | PLNT 5413 - Fall `r format(Sys.Date(),'%Y')`
output: pdf_document
link-citations: true
margin: 1in
fontsize: 12pt
geometry: "left=1in,right=1in,top=1in,bottom=1.25in"
header-includes:
- \usepackage{titling}
- \usepackage{lastpage}
- \usepackage{fancyhdr}
- \newcommand{\aspace}[1]{\vspace{#1\baselineskip}}
- \renewcommand{\headrulewidth}{0pt}
- \fancypagestyle{plain}{ \fancyhf{} \fancyfoot[C]{\aspace{2}Page \thepage~of~\pageref{LastPage}} }
- \pagestyle{fancy}
- \fancyhf{}
- \fancyfoot[C]{\small \thetitle \\ Data Science for Agriculture and Natural Resources\\PLNT 5413 - Fall `r format(Sys.Date(),'%Y')`\\ Page \thepage~of~\pageref{LastPage}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(agridat)
```

Load the `agridat` package and use the `diggle.cow` dataset with `ggplot()` to generate the following plots:

1. A scatterplot of weight over time

```{r prob1}
diggle.cow %>% 
  ggplot(aes(x = day, y = weight)) +
  geom_point()
```

2. A boxplot of weight for each day of measurement

```{r prob2}
diggle.cow %>% 
  ggplot(aes(x = day, y = weight, group = day)) +
  geom_boxplot()
```

3. A boxplot of the final weights for each treatment

```{r prob3}
diggle.cow %>% 
  filter(day==781) %>%
  ggplot(aes(x = interaction(iron,infect),
             y = weight,
             group = interaction(iron,infect))) +
  geom_boxplot()
```

4. A multi-panel line plot of weight over time for each animal within each treatment group

```{r prob4}
diggle.cow %>% 
  ggplot(aes(x = day, y = weight, group = animal)) +
  geom_line() +
  facet_grid(iron~infect)
```

5. A single-panel line plot of weight over time averaged within each treatment group and date of measurement

```{r prob5}
diggle.cow %>%
  group_by(iron,infect,day) %>%
  summarize(weight=mean(weight,na.rm=TRUE)) %>%
  ggplot(aes(x=day,y=weight,color=interaction(iron,infect))) +
    geom_line()
```
